---
title: "ESM 204 HW4"
author: "Tommy King"
date: "5/23/2022"
output: html_document
---

```{r setup, include=FALSE, message = FALSE, warning= FALSE}
knitr::opts_chunk$set(echo = FALSE, message= FALSE, warning = FALSE)

# loading packages 
library(tidyverse)
library(here)
library(dplyr)
library(janitor)

# The first, “damages.csv” provides new model estimates of the annual total damages from climate change at different levels of warming (in degrees C). The second dataset, “warming.csv” contains estimates of a baseline future climate trajectory (in degrees C) until 2100, and a second trajectory that adds a one-time pulse of CO2 today to the atmosphere. The pulse is 35 billion tons of carbon, which is roughly equal to annual global emissions. You can think of this as a “small” one-time pulse in carbon emissions.

# reading in the data and mutating it to show values in billions of dollars

damages_subset <- read_csv(here("damages.csv")) %>% 
  clean_names() %>% 
  mutate(damages = damages/1000000000)

warming_subset <- read_csv(here("warming (1).csv")) %>% 
  clean_names()

```

### **Question 1: Damage Equation**
```{r}
#Using damages.csv, estimate a quadratic damage function relating the dollar value of damages to the change in global mean temperature. Estimate a model with a zero intercept because damages by construction must equal zero when there is no climate change. Plot your estimated damage function, overlaid with a scatterplot of the underlying data.

#


#Squaring the warming variables. 
warming_new <- damages_subset$warming^2

# defining the damages quadratic function
damages_function <- lm(damages ~ 0 + warming + warming_new, data = damages_subset)

warming <- damages_subset$warming
damages_2 <- predict(damages_function)

damage_function <- function(x){
  damages_function$coefficients[1]*x+damages_function$coefficients[2]*x^2
}

```

#### **Figure 1: Damages Resulting from Climate Change**
```{r}
#graphing the function
ggplot(data = damages_subset, aes(x = warming, y = damages)) + 
  geom_point(color = "blue") + 
  geom_smooth(aes(x = warming, y = damages_2, color = "red")) + 
  labs(x = "Warming (Degrees Celsius)", y = 
         "Damages ($ billions)")
```

### **Question 2: Predicted Damages**
```{r}
# Use warming.csv and your estimated damage function to predict damages in each year under the baseline climate and the pulse scenario. Make four plots: (1) damages over time without the pulse, (2) damages over time with the pulse, (3) the difference in damages over time that arises from the pulse, and (4) the difference in damages over time from the pulse per ton of CO2 (you can assume that each ton of the pulse causes the same amount of damage).

# define the pulse 
pulse_tons <- 35 

# creating new variables for damages that reflect the warming variables from the warming subset

damages <- warming_subset %>% 
  mutate(damages_baseline = damage_function(warming_baseline)) %>% 
  mutate(damages_pulse = damage_function(warming_pulse)) %>% 
  mutate(damages_difference = damages_pulse - damages_baseline) %>% 
  mutate(damages_difference_tons = damages_difference/pulse_tons)

# creating graph number 1 for both scenarios and total damages

ggplot(data = damages, aes(x=year)) + 
  geom_smooth(aes(y= damages_baseline), color = "blue")+ 
  geom_smooth(aes(y=damages_pulse), color = "orange") + 
  labs(x = "Year", y = "Damages ($ billions)")

```


*Visualization Analysis: Difference between pulse and no pulse scenario appears to be trivial.* 

#### Figure 3: Absolute vs. Per Ton Damages Over Time 
```{r}
# creating graph showing the diffrences between absolute and per ton damages over time between the two scenarios. Facet wrapping two plots to show these differences.

baseline_graph <- ggplot(data= damages, 
                        aes(x = year, y = damages_difference)) + 
  geom_smooth(color = "red") + 
  labs(x = "Year", y = "Difference in Damages ($ billions)")

baseline_graph

# Creating second plot in tons 

tons_graph <- ggplot(data = damages, 
                     aes(x = year, y = damages_difference_tons)) + 
  geom_smooth(color = "purple") + 
  labs(x= "Year", y = "Difference in Damages ($ billions per ton")

tons_graph

```

**Figure 3 Analysis: Difference in damages rate over time moderately decreases, however, the overall damages increases exponentially from current day to 2100.**


### **Question 3: SCC Analysis**
```{r}
# The SCC is the present discounted value of the stream of future damages caused by one additional ton of CO2. The Obama Administration used a discount rate of 3% to discount damages. Recently, New York State used a discount rate of 2%. 

#Calculate and make a plot of the SCC (y-axis) against the discount rate (x-axis) for a reasonable range of discount rates. Explain the intuition for how the discount rate affects the SCC.

# Creating graph using a range of discount rates to see how the different rates affect the SCC values on the y-axis 

# First, must find total damages to 2100 from one ton of CO2 by adding together all of damages_difference_tons

total_damages_one_ton <- sum(damages$damages_difference_tons)

# now must create a function of NPV use disount rate, damages, time, and present time

discount_damages <- function(rate, damage, time, present_time){
  NPV <- damage/ (1 + (rate/100))^ (time- present_time) 
  return(NPV)
}

# now create a new data frame for different discount rates 

discount_table <- data_frame(rates = seq(0,10,0.5), SCC = seq(0,10,0.5))

# now use a for loop to find discounted damages up to 2100 for discount rates from 0 to 10 and sum to find total damages per discount rate

for(x in 1:length(discount_table$rates)){
  discount_table$SCC[x] <- sum(discount_damages(discount_table$rates[x], damages$damages_difference_tons, damages$year, 2022))
}

# now we plot the for loop for discount to visualize this process 

ggplot(data = discount_table) + 
  geom_point(aes(x = rates, y = SCC), color = "darkblue") + 
  labs(x = "Discount Rates (%)", y = "Social Cost of Carbon (SCC) ($/ton)")
```

***Figure 4: Artificially increasing the discount rate lowers the SCC exponentially. Thus, the proposed rates by the Obama Administration and NY state lead to SCC values between $50 and 100.***

### **Question 4: Ramsey Rule** 
### Figure 5: Ramsey Rule SCC evaluation
```{r}
# The National Academies of Sciences, Engineering, and Medicine advised the government in a 2017 report to use the Ramsey Rule when discounting within the SCC calculation:
 #r = ρ + ηg
#Using ρ = 0.001, η = 2, and g = 0.01, what is the SCC? Locate this point on your graph from above.

# define values 

p <- 0.001
n <- 2
g <- 0.01

# plug into equation 

ramsey_rule <- (p+n*g)*100

# plug into previous damages equation from above 
ramsey_discount <- sum(discount_damages(ramsey_rule, damages$damages_difference_tons, damages$year,  2022))

# Scc is about $71.35 in Ramsey Rule

#Now plot that value 

ggplot(data = discount_table) + 
  geom_point(aes(x = rates, y = SCC), color = "darkblue") + 
  geom_point(x = ramsey_rule, y = ramsey_discount, color = "red", size = 4) + 
  geom_vline(xintercept= 2.1, color = "darkblue", linetype = "dashed") + 
  geom_label(aes(x = 6.5, label = "SCC under Ramsey Rule = $71.35", y = 72), color = "darkblue", angle = 90) + 
  labs(x = "Discount Rates(%)", y = "Social Cost of Carbon ($/ton)")

```

NPV of future damages between current day and 2100 discounted at various levels with different discount rates (0-10% in increments of 0.5%). The red point shows the discount rate and accompanying SCC under the Ramsey Rule. 

The SCC at the proposed discount rate under Ramsey Rule is $`r round (ramsey_discount,2)`/ton. 

### **Question 5: Comparing Policy Considerations**

Question 5 asks to consider business as usual policy (Policy A) compared to strong action against climate change (Policy B) in which: 

***Policy A:*** Either warming will occur as in the “baseline” (i.e. “no-pulse”) dataset above (this happens with probability 0.5) or warming each year will be 1.5 times that in the “baseline” dataset (with probability 0.5).

***Policy B:*** Warming will continue until 2050 as in the “baseline” dataset, and then will stabilize at 1.29 degrees and stay that way forever.
```{r}
#5. Now suppose there are two possible climate policies that can be pursued. 


#Policy A is business as usual and Policy B is to take immediate and strong action on climate change. 

#Use these facts
# • If you undertake Policy A there are two possible outcomes. Either warming will occur as in the “baseline” (i.e. “no-pulse”) dataset above (this happens with probability 0.5) or warming each year will be 1.5 times that in the “baseline” dataset (with probability 0.5).

#• Under Policy B, warming will continue until 2050 as in the “baseline” dataset, and then will stabilize at 1.29 degrees and stay that way forever.

#• Society is risk neutral
#• Use a discount rate of 2%

#What is the expected present value of damages up to 2100 under Policy A? What is the expected present value of damages up to 2100 under Policy B? Suppose undertaking Policy A costs zero and undertaking Policy B costs X. 

# The total cost of a policy is the implementation cost plus expected damages. Make a plot of total cost of Policy A and total cost of Policy B against X. Suppose Congress wants to undertake the policy with the lower total cost. Use your graph to advise Congress.

# Estimate PV for the Policy A scenario, discount rate is 2% 

discount_policy_a <- 2

# create a new dataset with the damages and PV for both baseline and 1.5x warming 

damages_a <- damages %>% 
  select(year, warming_baseline, damages_baseline) %>% 
  mutate(damages_1.5 = damage_function(1.5*warming_baseline)) %>% 
  mutate(pv_baseline = discount_damages(discount_policy_a, damages_baseline, damages$year, 2022)) %>% 
  mutate(pv_1.5 = discount_damages(discount_policy_a, damages_1.5, damages$year, 2022))

# now calculating probabilities for both 1.5 and base 

probability_base <- 0.5 
probability_1.5 <- 0.5 

# now summing PV for each 

sum_base <- sum(damages_a$pv_baseline)
sum_1.5 <- sum(damages_a$pv_1.5)

# now scale by each probability 
ev_a <- (sum_base * probability_base) + (sum_1.5 * probability_1.5)


# Now moving to policy B, same process

damages_b <- damages %>% 
  select(year, warming_baseline) 

for(i in 30:79){ 
  damages_b$warming_baseline[i]=1.29
}

# now warming is locked at 1.29 after 2050

damages_b <- damages_b %>%  
  mutate(damages_b = damage_function(warming_baseline)) %>% 
  mutate(pv_b = discount_damages(discount_policy_a, damages_b, damages$year, 2022))

prob_b <- 1
sum_b <- sum(damages_b$pv_b)
ev_b <- (sum_b * prob_b)

```

- **Scenario A: PV of Damages under Policy A=** $`r round(ev_a, 0)` billion. 
- **Scenario B: PV of Damages under Policy B=** $`r round(ev_b, 0)` billion. 

